---
title: "College Analysis"
author: "Alexander Chen"
date: "2/22/2023"
output: 
  rmdformats::readthedown:
    number_sections: T
    code_folding: "hide"
    fig_width: 12
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = F}
#Knit default settings
knitr::opts_chunk$set(echo = T, fig.width=8, fig.height=4, warning = F)

#Options
options(scipen = 0, digits = 3, max.print = 500)

#Load required packages
if(!require("pacman")) install.packages("pacman")
pacman::p_load(devtools, tidyverse, data.table, lubridate, dplyr, rpart, rpart.plot, ggplot2, haven, magrittr, skimr, GGally, rmdformats, prettydoc, dplyr, DT, e1071, fmtr, Hmisc, readxl, lattice, readstata13)
```

## 1. For every survey, provide a succinct but comprehensive description of the nature of the data included.  Among the pieces of information I need:
## 1a. What is the source of the information
## 1b. Is the information available for all colleges? If not, which ones are excluded? Why?
## 1c. Is there any reason that makes you believe the variable definition is not consistent across colleges? Why?
## 2. For every survey, provide me with the dataset in stata format, complete with the variables' description included in the DICT file
## 3. Merge all surveys in one dataset, where each raw refers to one college

```{r read data}
directory.raw <- as.data.frame(read.csv("data/directory/hd2021.csv"))
directory.dict <- read_excel("data/directory/hd2021.xlsx", sheet = 3)

# Adding notes to directory.raw
directory.notes <- directory.dict$longDescription
directory.raw <- addNotes(directory.raw, directory.notes)
attributes(directory.raw)
save.dta13(directory.raw, "directory_raw.dta")






library(readstata13)
directory.dta <- read.dta13("directory_rawnotesExport.dta")
attributes(directory.dta)

# testing adding a note to var.
# Always three rows. Row 1 lists variable name, 
# row two lists content (if initialization column, is integer), row three lists note number

# TASKTWO: WRITE FUNCITON THAT TAKES A DATAFRAME AND LIST
# MATCHING APPLIES NOTE TO EVERY COLUMN IN DATAFRAME

# notes.dt <- data.frame(nrow = 3)[,-1]
# for(i in 1:ncol(directory.raw)){
#   varname <- names(directory.raw)[i]
#   merger <- data.frame(c(toString(varname), "note0", 1))
#   notes.dt <- cbind(notes.dt, merger)
#   merger <- data.frame(c(toString(varname), "note1", "this is the best note"))
#   notes.dt <- cbind(notes.dt, merger)
# }
# view(notes.dt)
# notes.list <- as.list.data.frame(notes.dt)
# attr(directory.dta, "expansion.fields") <- notes.list
# attributes(directory.dta)






directory.dta <- addNotes(directory.dta, noteDummy)
view(attr(directory.dta, "expansion.fields"))











# attr(directory.raw, "label") <- as.list(directory.dict$longDescription)

for (i in 1:ncol(directory.raw)){
  attr(directory.raw[,i], "expansion.fields") <- "test if label works"
}

attributes(directory.raw$UNITID)


match(c(1, 2, 3, 4, 5), c(6, 7, 8, 2, 10))


comment(directory.raw$UNITID) <- "uniiiiiID"


sampleVar <- 10
attr(sampleVar, "new attribute") <- "testing new attribute"
attributes(sampleVar)

comment(names(directory.raw)[1]) <- "comment on UNITID name"
comment(directory.raw$UNITID)


comment(directory.raw) <- "directoryComment"
library(lattice)
comment
comment(directory.raw$UNITID)


# TASKONE: ASSIGNING LABELS TO EACH VARIABLE
label(directory.raw) <- as.list(directory.dict$longDescription)
write_dta(directory.raw, "directory_raw.dta")



# check for missingness
view(apply(directory_dt, c(1, 2), isMissing, missingnessList = c("", " ")))

# Documentation
# IC2021 Educational offerings, organization, services and athletic associations	
# IC2021_AY	Student charges for academic year programs	
# IC2021_PY	Student charges by program (vocational programs)	
# FLAGS2021	Response status for all survey components	(regarding nature of responses. Such as imputation, special data reported in aggregate from multiple campuses, etc.)
# 
# EFFY2021 12-month unduplicated headcount by race/ethnicity, gender and level of student: 2020-21	
# EFFY2021_DIST	12-month unduplicated headcount by distance education status and student level: 2020-21	
# EFIA2021 12-month instructional activity: 2020-21	
# FLAGS2021	Response status for all survey components	
# 
# ADM2021	Admission considerations, applications, admissions, enrollees and test scores, fall 2021	
# FLAGS2021	Response status for all survey components	
# 
# EF2021A	Race/ethnicity, gender, attendance status, and level of student: Fall 2021	
# EF2021B	Age category, gender, attendance status, and level of student: Fall 2021	
# EF2021C	Residence and migration of first-time freshman: Fall 2021	
# EF2021D	Total entering class, retention rates, and student-to-faculty ratio: Fall 2021	
# EF2021A_DIST	Distance education status and level of student: Fall 2021	
# FLAGS2021	Response status for all survey components	
# 
# C2021_A	Awards/degrees conferred by program (6-digit CIP code), award level, race/ethnicity, and gender: July 1, 2020 to June 30, 2021	
# C2021_B	Number of students receiving awards/degrees, by race/ethnicity and gender: July 1, 2020 to June 30, 2021	
# C2021_C	Number of students receiving awards/degrees, by award level and by gender, race/ethnicity and age categories: July 1, 2020 to June 30, 2021	
# C2021DEP	Number of programs offered and number of programs offered via distance education, by award level: July 1, 2020 to June 30, 2021	
# FLAGS2021	Response status for all survey components	
# 
# SAL2021_IS	Number and salary outlays for full-time nonmedical instructional staff, by gender, and academic rank: Academic year 2021-22	
# SAL2021_NIS	Number and salary outlays for full-time nonmedical noninstructional staff, by gender, and academic rank: Academic year 2021-22	
# FLAGS2021	Response status for all survey components	
# 
# S2021_OC	Full- and part-time staff by occupational category, race/ethnicity, and gender: Fall 2021	
# S2021_SIS	Full-time instructional staff, by faculty and tenure status, and academic rank (Degree-granting institutions): Fall 2021	
# S2021_IS	Full-time instructional staff, by faculty and tenure status, academic rank, race/ethnicity, and gender (Degree-granting institutions): Fall 2021	
# S2021_NH	New hires by occupational category, race/ethnicity, and gender (Degree-granting institutions): Fall 2021	
# FLAGS2021	Response status for all survey components	
# 
# EAP2021	Number of staff by occupational category, faculty and tenure status: Fall 2021	
# FLAGS2021	Response status for all survey components	
# 
# F2021_F1A	Public institutions - GASB 34/35: Fiscal year 2021	
# F2021_F2	Private not-for-profit institutions or Public institutions using FASB: Fiscal year 2021	
# F2021_F3	Private for-profit institutions: Fiscal year 2021	
# FLAGS2021	Response status for all survey components	
# 
# SFA2021	Student financial aid and net price: 2020-21	
# SFAV2021	Military Servicemembers and Veteran's Benefits: 2020-21	
# FLAGS2021	Response status for all survey components	
# 
# GR2021	Graduation rate data, 150% of normal time to complete - cohort year 2015 (4-year) and cohort year 2018 (2-year) institutions	
# GR2021_L2	Graduation rate data, 150% of normal time to complete - cohort year 2018 (less-than-2-year institutions)	
# GR2021_PELL_SSL	Graduation rate data for Pell Grant and Subsidized Stafford Loan recipients, 150% of normal time to complete - cohort year 2015 (4-year) and cohort year 2018 (less-than-4-year) institutions	
# GR200_21	Graduation rate data, 200% of normal time to complete - cohort year 2013 (4-year) and cohort year 2017 (less-than-4-year) institutions	
# FLAGS2021	Response status for all survey components	
# 
# OM2021	Award and enrollment data at four, six and eight years of entering degree/certificate-seeking undergraduate cohorts from 2013-14 at degree-granting institutions, by Pell status	
# FLAGS2021	Response status for all survey components	
# 
# AL2021	Academic Libraries: Fiscal year 2021	
# FLAGS2021	Response status for all survey components	

gradRate_flags_dt <- as.data.frame(read.csv("data/gradRates/flags2021.csv"))

summary(gradRate_flags_dt)

label(data) = lapply(names(data), function(x) var.labels[match(x, names(var.labels))])

# Import list of all US Colleges
```

```{r functions}
# Prerequisites
# The length of datatable and noteslist must be the same.
# In addition, datatable must contain STATA exportable attributes
# created with readstata13.
addNotes <- function(datatable, notesList){
  if(ncol(datatable) != length(notesList)){
    print("notes failed")
    return(NULL)
  }
  else{
    notes.dt <- data.frame(nrow = 3)[,-1]
    for(i in 1:ncol(datatable)){
      varname <- names(datatable)[i]
      varnote <- notesList[i]
      merger <- data.frame(c(toString(varname), "note0", 1))
      notes.dt <- cbind(notes.dt, merger)
      merger <- data.frame(c(toString(varname), "note1", varnote))
      notes.dt <- cbind(notes.dt, merger)
    }
    notes.list <- as.list.data.frame(notes.dt)
    attr(datatable, "expansion.fields") <- notes.list
  }
  return (datatable)
}

isMissing <- function(data, missingnessList){
  if(data %in% as.list(missingnessList)){
    return (TRUE)
  }
  return (FALSE)
}
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
